---
title: "HW1"
author: "顧以恩 RE6131066"
format: 
  pdf:
    documentclass: ctexart
    number-sections: true
    mainfont: "SimSun"  
    latex-engine: xelatex
editor: visual
warning: False
---

## Titanic dataset

#### Import packages and read data

```{r}
#| label: import packages and read data

#install.packages("titanic")
library(titanic)
library(dplyr)
library(ggplot2)
library(GGally)


data(titanic_train)
head(titanic_train)
```

#### Variables describtion

```{r}
#| label: summary data

summary(titanic_train)

```

#### 

-   **Survival**: Survival status (0 = No 38%, 1 = Yes 62%)

<!-- -->

-   **Pclass**: Ticket class (Ordinal variable: 1 = 1st, 2 = 2nd, 3 = 3rd)

-   **Sex**: Gender (Nominal variable: male = Male 65%, female = Female 35%)

-   **Age**: Age (Continuous variable: 0.42 - 80, 177 missing values)

-   **SibSp**: Number of siblings or spouses aboard the ship (Ordinal variable: 0 - 8)

-   **Parch**: Number of parents or children aboard the ship (Ordinal variable: 0 - 6)

-   **Ticket**: Ticket number

-   **Fare**: Ticket price (Continuous variable: 0 - 512, no missing values)

-   **Cabin**: Cabin number

-   **Embarked**: Port of embarkation (C = Cherbourg 19%, Q = Queenstown 9%, S = Southampton 72%, 2 missing values)

#### Transformation

```{r}
#| label: transform

# Sex should be binary categorical variable
# Therefore transform it into factor
print(length(table(titanic_train$Sex)))
titanic_train$Sex = as.factor(titanic_train$Sex)

#Embarked is also categorical variable, transform it into factor
print(length(table(titanic_train$Embarked)))
titanic_train$Embarked = as.factor(titanic_train$Embarked)
titanic_train = titanic_train %>% mutate(Survived = as.factor(Survived),
                                         Sex = as.factor(Sex),
                                         #SibSp = as.factor(SibSp),
                                         #Parch = as.factor(Parch),
                                         Embarked = ifelse(Embarked=="",NA,Embarked) %>% as.factor()
                                         )


# Name, Ticket and Cabin variables are all charactors
head(titanic_train$Name,5)
head(titanic_train$Ticket,5)
head(titanic_train$Cabin,5)

titanic_train = titanic_train %>% select(-c(PassengerId, Name, Ticket, Cabin))

summary(titanic_train)
```

#### Visuallization

```{r ggpairs_plot, fig.width=12, fig.height=12, out.width="100%"}


ggpairs(
  titanic_train,
  title = "Pairwise Relationships in Titanic Dataset",
  lower = list(continuous = wrap("smooth", method = "lm", se = FALSE)), # 下三角加入回歸線
  upper = list(continuous = wrap("cor", size = 4)), # 上三角顯示相關係數
  diag = list(continuous = "barDiag"), # 對角線繪製直方圖
  mapping = aes(color = Survived) # 顏色區分存活
) +
  theme_minimal()


```

a)  Survival Analysis (Survived) More people did not survive (Survived = 0, shown in red). The survival rate seems to vary significantly with Sex and Pclass.

b)  Passenger Class (Pclass) First-class (Pclass = 1) passengers had a higher fare and survival rate. Third-class (Pclass = 3) passengers had the lowest survival rate.

c)  Gender Influence (Sex) Female passengers had a higher survival rate than males. Males had a lower chance of survival, as indicated by the red proportion in the Sex vs. Survived plots.

d)  Age Distribution (Age) Young children had a higher survival rate. The majority of passengers were adults. There is a negative correlation between Age and Survival (older passengers had a lower chance of survival).

e)  Family Members (SibSp and Parch) Having more siblings/spouses (SibSp) or parents/children (Parch) slightly increases survival probability, but very high numbers reduce survival chances. Large families had a lower survival rate compared to single passengers.

f)  Fare and Survival (Fare) Higher Fare is positively correlated with survival (Corr = 0.538\*\*\*), indicating that wealthier passengers had better chances. This aligns with the observation that first-class passengers survived more.

g)  Embarkation Port (Embarked) Most passengers embarked from Southampton (S). Higher survival rates are observed among passengers from Cherbourg (C), likely due to a higher proportion of first-class passengers.

#### Appendix

```{r}
ggplot(titanic_train, aes(x = Age)) +
  geom_histogram(binwidth = 5, fill = "steelblue", color = "black") +
  labs(title = "Age Distribution", x = "Age", y = "Count")
```

```{r}

ggplot(titanic_train, aes(x = Fare, fill = as.factor(Survived))) +
  geom_histogram(binwidth = 20, fill = "lightcoral", color = "black") +
  labs(title = "Fare Distribution", x = "Fare", y = "Count")
```

```{r}
ggplot(titanic_train, aes(x = as.factor(Survived), y = Age, fill = as.factor(Survived))) +
  geom_boxplot() +
  labs(title = "Age Distribution by Survival", x = "Survived", y = "Age") +
  theme_minimal()
```

```{r}
ggplot(titanic_train, aes(x = as.factor(Survived), y = Fare, fill = as.factor(Survived))) +
  geom_boxplot() +
  labs(title = "Fare Distribution by Survival", x = "Survived", y = "Fare") +
  theme_minimal()
```

```{r}
ggplot(titanic_train, aes(x = Pclass, fill = as.factor(Survived))) +
  geom_bar(position = "fill") +
  labs(title = "Survival Rate by Passenger Class", x = "Passenger Class", y = "Proportion") +
  theme_minimal()
```

```{r}
ggplot(titanic_train, aes(x = Sex, fill = as.factor(Survived))) +
  geom_bar(position = "fill") +
  labs(title = "Survival Rate by Gender", x = "Sex", y = "Proportion") +
  theme_minimal()

```

```{r}
library(ggcorrplot)

cor_matrix <- cor(titanic_train[, sapply(titanic_train, is.numeric)], use = "complete.obs")
ggcorrplot(cor_matrix, lab = TRUE)

```
